Simulation Learning in Computer Hex

### 2011-08-24 thesis/thesis.tex
Needs editing and some comments on what to improve, I'm mostly out of ideas.

\subsection{Fitness Evaluation}\label{fitness}
The ES population is initialized with $\lambda$ individuals, all with uniform random weights in the range $[0, 100]$. In practice, weights are not initialized until a pattern lookup is performed (a form of lazy evaluation).

Every generation, each individual plays $n$ games against another random individual. Roughly half the games are played as black, half as white. The swap-safe rule was used to play the first move to further equalize the first-mover advantage. The winner of a game receives 1 point added to their fitness, the loser has 1 point deducted from their fitness.

\section{Results and Conclusions}\label{results}
To test the effects of learning policy weights, the evolution process was ran for 41 generations. This took approximately 2 days using a single core of an Intel Xeon CPU running at 2.80GHz. Although evolution strategies is easy to parallelize by running multiple evaluations simultaneously, the current implementation is single-threaded. After training was completed, the weights of top 4 individuals of the 41st generation were blended in proportion to their fitness to form a single individual. This final individual was used to bias the simulation policy and compared to the uniform local bias, uniform local bias with a $\frac{5}{6}$th probability of playing locally. The control consisted of uniform random playouts. All players used a slightly modified UCT selection policy that included the variance term suggested by Gelly 2006\cite{gelly2006exploration}. A pure update policy was used, only the nodes directly in the path traversed by UCT were updated with the result of a simulation.

@@

>>>>

\subsection{Fitness Evaluation}\label{fitness}
The ES population is initialized with $\lambda$ individuals, with uniform random weights in the range $[0, 100]$. In practice, weights are not initialized until a pattern lookup is performed (lazy lookup). Every generation, each individual plays $n$ complete games against another random individual. At the beginning of the game, one of the two individuals is randomly chosen to play \black, the other to play \white. The swap-safe rule was used to play the first move to further equalize first-mover advantage. The winner of a game receives 1 point added to their fitness, the loser has 1 point deducted from their fitness.

\section{Results and Conclusions}\label{results}
To test the effects of learning policy weights, the evolution process was ran for 41 generations. This took approximately 2 days using a single core of an Intel Xeon CPU running at 2.80GHz. Although evolution strategies is easy to parallelize by running multiple evaluations simultaneously, the current implementation is single-threaded. After training was completed, the 41st generation was used to form a single individual. This final individual was used to bias the simulation policy and compared to the uniform local bias, uniform local bias with a $\frac{5}{6}$th probability of playing locally. The control consisted of uniform random playouts. All players used a slightly modified UCT selection policy that included the variance term suggested by Gelly 2006\cite{gelly2006exploration}. A pure update policy was used, only the nodes directly in the path traversed by UCT were updated with the result of a simulation.

@@

Instead of just using the best individual from the last generation, the weight vector used is a linear combination of the weights from all individuals in generation 41. The weights are blended based on relative fitness, so the final weight vector will most reflect the top performer of generation 41, but also includes the weights off all individuals. In some respect all individuals have ``survived'' 41 generations of recombination and mutation, and so were included in the final learned vector.
